---
title: "Practical machine learning"
author: "Julian Flowers"
date: "6 March 2016"
output: html_document
---
## Summary

This exercise presents summary and analysis of the weight lifting exercises dataset as part of the Coursera Practical Machine Learning course.  

The data is [here](http://groupware.les.inf.puc-rio.br/har). It consists of a series of measures on 6 healthy volunteers of 10 repetitions of Unilateral Dumbell Curls (UDC) performed in 5 different ways of which one was correct. The objective of the analysis is to predict which form of UDC each of 20 test subjects performed.

First we need to downoad and examine the data.

## Downloads
```{r, cache = TRUE, message=FALSE}

train <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("NA","#DIV/0!"))

test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings = c("NA","#DIV/0!"))
dim (train)
```


## Extract the sensor data

We need to limit the dataset to the measure variables only, and we will exclude those variables with high proportions of missing data. 

```{r}
sensordata <- grep(pattern = "_belt|_arm|_dumbbell|_forearm", names(train))
data <- train[,c(sensordata, 160)]
datatest <- test[,c(sensordata, 160)]

nadata<- apply(data, 2, function(x) mean(is.na(x)))
t <- which(nadata >0)
train1 <- data[,-t]

nadatatest<- apply(datatest, 2, function(x) mean(is.na(x)))
t <- which(nadatatest >0)
test1 <- datatest[,-t]

dim(train1); dim(test1)
```

## Loading necessary packages
```{r}
library(caret)
library(randomForest)
library(corrplot)
```

## Setting up training and test sets 

```{r}
## partition data
set.seed(123)
train <- createDataPartition(train1$classe, p=0.6, list = FALSE)
inTrain <- train1[train,]
inTest <- train1[-train,]
dim(inTrain); dim(inTest)
```

## Examine relationships between variables

```{r}
library(corrplot)
cor <- cor(inTrain[, -53])
corrplot(cor, order = "FPC", method = "color", type = "lower", tl.cex = 0.7, tl.col = rgb(0, 0, 0))

```

## Tree approach

```{r}
modtree <- train(classe~., data = inTrain, method = "rpart")

plot(modtree$finalModel)
text(modtree$finalModel, pretty = 1)
predtree <- predict(modtree, inTest)
confusionMatrix(predtree, inTest$classe)
```

Overall accuracy is poor at 0.51.


```{r}
set.seed(321)
modrf <-randomForest(classe ~., data = inTrain, ntree = 500)
modrf
varImpPlot(modrf, cex = 0.5)

```



```{r}
pred <- predict(modrf, inTest)
confusionMatrix(pred, inTest$classe)

```



```{r}
pred1 <- predict(modrf, test1)
pred1
```

